---
title: "Stat 251 HW 6"
author: "Neil Thompson"
date: "3/15/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#1a
```{r 1a}
lambda <- 1000
tau <- 200
sigma <- 40
data <- c(1010, 1000, 950, 1050)
n <- length(data)
lambda1 <- (tau^2*sum(data) + sigma^2*lambda)/(tau^2*n + sigma^2)
tau1 <- sqrt(sigma^2*tau^2/(tau^2*n + sigma^2))
lambda1
tau1
```
Posterior distribution: 
$\mu |$data$, \sigma^2 \sim \mathcal{N}(1002.475, 19.90074^2)$

#1b
```{r 1b}
qnorm(c(0.025, 0.975), lambda1, tau1)
```
There is a 95% chance that $\mu$ is between 963.47 and 1041.48.

#1c
```{r 1c}
lamZeta <- 2203 - 0.835*lambda1
tauZeta <- 0.835*tau1
lamZeta
tauZeta
```
Posterior distribution: 
$\zeta |$data$, \sigma^2 \sim \mathcal{N}(1365.93, 16.617^2)$

#1d
```{r 1d}
qnorm(c(0.025, 0.975), lamZeta, tauZeta)
```
There is a 95% chance that $\mu$ is between 1333.364 and 1398.502.

#2a
```{r 2a}
lambdaS <- 1000
tauS <- 200
sigmaS <- 40
dataS <- c(940, 1040, 910, 990)
nS <- length(dataS)
lambdaS1 <- (tauS^2*sum(dataS) + sigmaS^2*lambdaS)/(tauS^2*nS + sigmaS^2)
tauS1 <- sqrt(sigmaS^2*tauS^2/(tauS^2*nS + sigmaS^2))
lambdaS1
tauS1
```
Posterior distribution: 
$\mu_S |$data$, \sigma^2 \sim \mathcal{N}(970.297, 19.90074^2)$

#2b
```{r 2b}
lamZetaS <- 2203 - 0.835*lambdaS1
tauZetaS <- 0.835*tauS1
lamZetaS
tauZetaS
```
Posterior distribution: 
$\zeta_S |$data$, \sigma^2 \sim \mathcal{N}(1392.8, 16.617^2)$

#2c
```{r 2c}
lam_d <- lamZeta - lamZetaS
tau_d <- sqrt(tauZeta^2 + tauZetaS^2)
lam_d
tau_d
```
Posterior distribution: 
$d |$data$, \sigma^2 \sim \mathcal{N}(-26.869, 23.5^2)$

#2d
```{r 2d}
qnorm(c(0.05, 0.95), lam_d, tau_d)
```
There is a 90% chance that the mean difference between the calibrated dates for Houhora and Shag River Mouth is between -65.523 and 11.786.

#2e
```{r 2e}
curve(dnorm(x, lam_d, tau_d), from = -100, to = 46)
abline(v = 0)
```
Definitely not. Our mean calibrated date for Shag River was 1393 and Houhora was 1366. There is clearly not going to be evidence to say that Shag River was inhabited first. That would be implying that the area in this plot to the left of the vertical line is insignificant.