---
title: 'Stat 251: HW 4'
author: "Neil Thompson"
date: "2/25/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
1a. 
```{r}
thetaM <- 67/1260
thetaW <- 173/1700
alpha <- 0.5
beta <- 3
curve(dbeta(x,alpha,beta))
pbeta(0.1,alpha,beta)
```
Prior probability that the proportion of skipped questions is less that 10% is 0.5546.

1b.
```{r}
alphaW <- 0.5 + 173
betaW <- 3 + 1700 - 173
alphaM <- 0.5 + 67
betaM <- 3 + 1260 - 67
alphaW
betaW
alphaM
betaM
```
$\theta_W|data \sim$ Beta(173.5, 1530)

$\theta_M|data \sim$ Beta(67.5, 1196)

1c. 
```{r}
pbeta(0.1,alphaW,betaW)
pbeta(0.1,alphaM,betaM)
```
Posterior probability that women will skip less than 10% is 0.4081.
Posterior probability that men will skip less than 10% is approximately 1.
For women, this probability is lower than that of the prior distribution, but for men this one is much higher.

1d. 
```{r}
curve(dbeta(x,alphaW,betaW))
curve(dbeta(x,alphaM,betaM))
thetaW.samp <- rbeta(1e6, alphaW, betaW)
thetaM.samp <- rbeta(1e6, alphaM, betaM)
diff.samp <- thetaW.samp - thetaM.samp
plot(density(diff.samp), xlab=expression(thetaW - thetaM), ylab="posterior
density", main="Posterior Difference between thetaW and thetaM")
mean(diff.samp<0)
```
Yes, women will almost surely skip more questions than men. The probability that men skip more than women is only 1 in 1000000.

2a.
```{r}
sdP <- 2.2
lambda <- sdP^2
lambda
```
$\lambda = 4.84$

2b.
```{r}
ppois(3, 1)
ppois(3, 10)
ppois(3, 100)
```

2c.
```{r}
plot(0:10, dpois(0:10, 1), pch = 19, xlab = "X", ylab = "Probability")
plot(0:30, dpois(0:30, 10), pch = 19, xlab = "X", ylab = "Probability")
plot(0:200, dpois(0:200, 100), pch = 19, xlab = "X", ylab = "Probability")
```
The $\lambda = 1$ distribution is not shaped the same as the other two distributions, which are bell shaped. Instead it has its highest values at $X = 0$ and $X = 1$. All three, naturally, range from 0 to infinity and are discrete (as are all Poisson distributions). The higher the $\lambda$ is, the larger the spread and the higher the center.

2d.
```{r}
qpois(0.95, 4)
```
The 95th percentile of Pois(4) is 8.

2e.
```{r}
ppois(8, 4)
```
It is higher than 95 because the distribution is discrete. 8 is the first value with a quantile above the 95th percentile and that is why 2d. gave us such an answer.

3a. 
```{r}
curve(dgamma(x, 2, 3), xlim = c(0, 6))
curve(dgamma(x, 6, 3), xlim = c(0, 6))
```
We see that as the value of $\gamma$ increases we see less right-skewedness and a larger mean and spread.

3b.
```{r}
curve(dgamma(x, 5, 2), xlim = c(0, 6))
curve(dgamma(x, 5, 9), xlim = c(0, 6))
```
We see that as the value of $\phi$ increases, both the mean and spread decrease but the shape remains the same.

3c.
```{r}
pgamma(5,2,1.5) - pgamma(2, 2, 1.5)
```
$P(2\leq\lambda\leq5|\gamma=2,\phi=1.5) = 0.1944$

3d.
```{r}
qgamma(0.025, 2, 1.5)
qgamma(0.975, 2, 1.5)
```
95% probability interval: (0.1615,3.714)







